---
title: "Access, clean and plot APRA Annual Fund Level Superannuation Statistics"
output: html_notebook
---

**NOTE: YOU SHOULD RUN ALL THE CHUNKS BEFORE TRYING TO PREVIEW A NOTEBOOK**

[INSERT NARRATIVE: what does this notebook examine/show ] 

## Access 

- To be completed by you

- Narrative would go here

```{r Chunk1_AccessData, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Insert libraries and prelim code here

```

## Tidy

- To be completed (by you)

- Description of what tidying is required would go here

- Insert code chunk that tidies data (from script.R)

```{r Chunk2_TidyData, echo=TRUE, message=FALSE, warning=FALSE}



```

## Summarise

- To be completed by you

- Insert code chunk and code that produces summary table (from script.R)

```{r Chunk3_SummariseData}

```

- Inspect `pander::pandoc.table` fundction that renders the `data.frame` as a table object
#### Super fund *mean* characteristics table
```{r Chunk4_SummaryTable, echo=FALSE, results='asis' }
pandoc.table(summary_table, 
             big.mark = " ", 
             digits = c(0,0,2,2,2,3,3,3),
             split.tables = Inf)
```

## Let's make some plots

- To be completed by you

- Insert code chunks that generate and display plots (from script.R)




