---
title: "Plotly"
author: "Marco Hatt"
date: "9 November 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r Access packages, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Libraries
packages <- c("gapminder","tidyverse","readxl","kableExtra","PCcharts","plotly","broom")

for (i in packages) { 
  if(!(i %in% rownames(installed.packages()))) install.packages(i)
}  

library(tidyverse)
library(readxl)
library(kableExtra) 
library(PCcharts)
library(plotly)
library(broom)
library(gapminder)
options(scipen = 999) 

```

## Interactive graphics with plotly

- Impressive

- Informative

- Easy


## From ggplot2 to plotly

**ggplot2**

- More features and flexibility
- Cannot do interactive graphics


**plotly**

- Interactive plots, great for websites and dashboards
- Also useful for exploring your data

  
## Example of an interactive plot

```{r plotly_fancy, echo=FALSE, message=FALSE, warning=FALSE}
gapminder %>%
  plot_ly(
    x = ~gdpPercap, 
    y = ~lifeExp, 
    size = ~pop, 
    color = ~continent, 
    frame = ~year, 
    text = ~country, 
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  ) %>%
  layout(
    xaxis = list(
      type = "log"
    )
  )
```


## It's this easy
**Tidy data**
```{r plotly_fancy_code, echo=FALSE, message=FALSE, warning=FALSE}
head(gapminder)
```
**Code**

gapminder %>%
  plot_ly(x = ~gdpPercap, y = ~lifeExp, size = ~pop, color = ~continent, frame = ~year, text = ~country, 
    hoverinfo = "text",type = 'scatter',mode = 'markers') %>%
  layout(xaxis = list(type = "log"))



```{r Download data, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}
# Download the APRA data if it is not there
if("2018-AFLSS-201706.xlsx" %in% list.files() == FALSE) { 
   link <- "https://www.apra.gov.au/sites/default/files/2018-AFLSS-201706.xlsx"
   download.file(link, "2018-AFLSS-201706.xlsx", mode = "wb")
   }

# Import Table 2
table2 <- read_xlsx("2018-AFLSS-201706.xlsx", sheet = "Table 2", range = "A5:AA206", col_names = TRUE, trim_ws = TRUE) 
table2 <- table2 %>% 
   filter(!is.na(`Fund name`) )
# Remove carriage return and line feed from colnames
colnames(table2) <- gsub("\r\n", " ", colnames(table2) )

# Import Table 3
table3 <- read_xlsx("2018-AFLSS-201706.xlsx", sheet = "Table 3", range = "A5:CE206", col_names = TRUE, trim_ws = TRUE)
table3 <- table3 %>%
   select(`Fund name`, `Total administration and operating expenses`, `Total Investment expenses`, `Cash flow adjusted net assets`) %>% 
   filter(!is.na(`Fund name`) )

# Join tables 2 and 3
apra <- table2 %>% 
   left_join(table3, by = "Fund name")

rm(table2, table3)
```

```{r TidyData, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}

apra <- apra %>% 
# Select required variables
   select(`Fund name`, `Fund type`, `Total assets`, `Number of MySuper products authorised`, 
          `Total number of member accounts`, `Average member account balance` ,
          `Total administration and operating expenses`, `Total Investment expenses` , `Cash flow adjusted net assets`, 
          `Proportion of total assets in default or MySuper strategy`,
          `One-year rate of return`, `Five-year rate of return`, `Ten-year rate of return` ) %>% 
# Convert vars that we want to be numeric
   mutate_at(.tbl = .,.vars = 3:length(.),.funs = as.numeric ) %>% 
# Turn proportions into percentages
   mutate_at(.tbl = .,.vars = 10:length(.),.funs = funs( .*100) ) %>% 
# Tidy fund type variable
   mutate(`Fund type` = case_when(`Fund type` == "Public Sector" ~ "Public sector" ,
                                `Fund type` == "Retail - ERF" ~ "Retail" , 
                                TRUE ~ `Fund type`) ) %>% 
   mutate(`Fund type` = fct_relevel(as_factor(`Fund type`), c("Industry", "Retail", "Corporate", "Public sector") ) ) %>% 
# Calculate operating and invesetment expenses ratios
   mutate(`Operating expense ratio` = `Total administration and operating expenses`/ `Cash flow adjusted net assets`*100 , 
          `Investment expenses ratio` = `Total Investment expenses`/ `Cash flow adjusted net assets`*100 ) %>% 
# Remove unnecessary variables
   select( -c("Total administration and operating expenses" , "Total Investment expenses" , "Cash flow adjusted net assets") )

#head(apra)

apra <- apra %>%
   filter(!is.na(`Total assets`) & `Total assets` != 0 ) %>%
   filter(!is.na(`Total number of member accounts`) & `Total number of member accounts` != 0 ) %>% 
# Assets are initially in $'000s
   mutate(`Total assets` = `Total assets`*1000) 
   
```


```{r OpExpPlot, echo=FALSE, message=FALSE, warning=FALSE}

Plot1_gg <- apra %>%
# Note: Total assets are squared here to emphasise differences in fund size
   mutate(`Total assets` = `Total assets`^2 ) %>% 
   ggplot( aes(x = `Operating expense ratio`, y = `Ten-year rate of return`) ) +
   geom_point( aes(size = `Total assets`, colour = `Fund type`), shape = 1, stroke=1.1) +
   geom_smooth()+
   PC.theme.scat() +
   scale_colour_manual(values = c(PC_green, PC_orange, PC_purple, PC_yellow) ) +
   scale_x_continuous( limits=c(0, 1) ) +
   guides(size=FALSE) 

```


## Two approaches
1. `ggplotly` function
  - Takes a `ggplot` object and returns a plotly object
  - Doesn't pick up all customisation in `ggplot`, 
  e.g. legend position
  - Easy way to add interactivity to `ggplot`
  
2. `plot_ly` interface
  - Faster (computationally) than `ggplotly`
  - Less features than `ggplot`


## ggplotly
- With ggplotly can take this static ggplot and make it interactive
```{r ggplot_static, echo=FALSE, message=FALSE, warning=FALSE}
Plot1_gg
```

## ggplotly
```{r ggplotly, echo=TRUE, message=FALSE, warning=FALSE}
ggplotly(Plot1_gg)
```


## Plotly basics 

- `ggplot` syntax: 
`ggplot(data,mapping=aes()) + geom_bar() ...`

- `plot_ly` syntax: 
 Borrows semantics from **dplyr** and **tidyr** packages

- Attributes defined via `plot_ly` which sets 'global' attributes that 
  are carried onto subsequent traces
  
- Example:
```{r plotly_101, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}

plot_ly(economics, x = ~date, color = I("black")) %>%
 add_lines(y = ~uempmed) %>%
 add_lines(y = ~psavert, color = I("red"))

```

##

```{r plotly_101_output, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}

plot_ly(economics, x = ~date, color = I("black")) %>%
 add_lines(y = ~uempmed) %>%
 add_lines(y = ~psavert, color = I("red"))

```

## Exercises

Using the APRA dataset

1. Using a box plot, identify the outlying funds by their
rate of return. Colour by fund type.
* hint: use `type = "box"`

2. Using a scatter plot, identify the outlyers in terms of 
operating expense ratio and investment expense ratio.
* hint: use `type = "scatter"`

`plot_ly(data = data.frame(), ..., color, text, type)`

- Remember the `~` e.g. x=~'variable name'

## Exercise answer 1: box plot
```{r plotly_basic, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE}

plot_ly(data=apra, x = ~`Ten-year rate of return`, 
        color = ~`Fund type`, text = ~`Fund name`, type = "box")

```

## Exercise answer 2: scatter plot

```{r plotly_basic2, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE}

plot_ly(data=apra, x = ~`Operating expense ratio`, 
        y = ~`Investment expenses ratio`, type = 'scatter', 
        mode = 'markers',text = ~`Fund name`, color = ~`Fund type`)
```

## subplot 

-The `subplot` function provides a flexible interface for merging multiple plotly objects into a single object

-A bit like `facet_wrap` and `facet_grid` from `ggplot`, e.g.

```{r facet_wrap, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}
apra %>% select(`Fund type`,`Operating expense ratio`,`Investment expenses ratio`,`Proportion of total assets in default or MySuper strategy`,`Ten-year rate of return`,`Total assets`) %>% 
  mutate(`Total assets` = `Total assets`^2) %>% 
  gather(key="Variable",value="value",-c(`Fund type`,`Ten-year rate of return`,`Total assets`)) %>% 
  ggplot(aes(y=`Ten-year rate of return`, x=value , colour=`Fund type`, size=`Total assets`)) + geom_point() +
  PC.theme.scat() +  
  facet_wrap(~Variable,scales="free")   +
  xlab("") +
  scale_colour_manual(values=PC.palette(n=4))
```

## subplot 

```{r subplot, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE, results='hide',size='tiny'}

p1 <- apra %>% 
  plot_ly(x = ~`Investment expenses ratio`, y = ~`Ten-year rate of return`, 
          type = 'scatter', mode = 'markers',text = ~`Fund name`, 
          color = ~`Fund type`,size = ~`Total assets`^2,
          legendgroup= ~`Fund type`,showlegend = F)

p2 <- apra %>% 
  plot_ly(x = ~`Operating expense ratio`, y = ~`Ten-year rate of return`, 
          type = 'scatter', mode = 'markers',text = ~`Fund name`, 
          color = ~`Fund type`,size = ~`Total assets`^2,
          legendgroup= ~`Fund type`,showlegend = F)
 
p3 <- apra %>% 
  plot_ly(x = ~`Proportion of total assets in default or MySuper strategy`, 
          y = ~`Ten-year rate of return`, type = 'scatter', 
          mode = 'markers',text = ~`Fund name`, color = ~`Fund type`,
          size = ~`Total assets`^2,
          legendgroup= ~`Fund type`,showlegend = T)

```



## subplot

```{r subplot_function, echo=TRUE, message=FALSE, warning=FALSE, eval=TRUE, results='hide'}

apra.range<- apra %>% 
  select(`Operating expense ratio`,`Ten-year rate of return`) %>% 
  na.omit()

min.y <- floor(round(min(apra.range$`Ten-year rate of return`),1))
max.y <- ceiling(round(max(apra.range$`Ten-year rate of return`),1))

subplot(p1,p2,p3,titleY = TRUE,titleX=TRUE,shareY = TRUE) %>% 
  layout(yaxis = list(range = c(min.y,max.y)))

```


## subplot 

```{r subplot_output, echo=FALSE, message=FALSE, warning=FALSE}

subplot(p1,p2,p3,titleY = TRUE,titleX=TRUE,shareY = TRUE) %>% 
  layout(yaxis = list(range = c(min.y,max.y)))

```

## Drop down menus
- What if you wanted to have a drop down box instead of (or in addition to) the legend filter

```{r filter_legend, echo=TRUE, message=FALSE, warning=FALSE}
plot_ly(apra, x = ~`Operating expense ratio`, y = ~`Ten-year rate of return`, 
        type = 'scatter', mode = 'markers',text = ~`Fund name`,
        size = ~`Total assets`^2,color= ~`Fund type`)
```

## Drop down menus

```{r filter_box, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE, results='hide'}
plot_ly(apra, x = ~`Operating expense ratio`, 
        y = ~`Ten-year rate of return`, 
        type = 'scatter', mode = 'markers',text = ~`Fund name`,
        size = ~`Total assets`^2,
        transforms = list(
          list( 
            type = 'filter',
            target = ~`Fund type`,
            operation = '=',
            value = ~`Fund type`[1]))) %>% 


```

## Drop down menus


```{r filter_box_layout, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE, results='hide'}
  layout(
    updatemenus = list(
      list(
        type = 'dropdown',
        active = 1,
        buttons = lapply(unique(apra$`Fund type`), function (x) {
          list(method = "restyle",
               args = list("transforms[0].value",x),
               label = x)} ) )),
    xaxis = list(range = c(min.x,max.x)),
    yaxis = list(range = c(min.y,max.y)))

```
## Drop down menus

```{r filter_box_output, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}


apra.range<- apra %>% select(`Operating expense ratio`,`Ten-year rate of return`) %>% na.omit()
min.x <- floor(round(min(apra.range$`Operating expense ratio`),1))
max.x <- ceiling(round(max(apra.range$`Operating expense ratio`),1))
min.y <- floor(round(min(apra.range$`Ten-year rate of return`),1))
max.y <- ceiling(round(max(apra.range$`Ten-year rate of return`),1))


  p <- plot_ly(apra, x = ~`Operating expense ratio`, y = ~`Ten-year rate of return`, 
          type = 'scatter', mode = 'markers',text = ~`Fund name`,size = ~`Total assets`^2,
          transforms = list(
            list( 
              type = 'filter',
              target = ~`Fund type`,
              operation = '=',
              value = ~`Fund type`[1]))) %>% 
  layout(
  updatemenus = list(
    list(
      type = 'dropdown',
      active = 1,
      buttons = lapply(unique(apra$`Fund type`), function (x) {
        list(method = "restyle",
             args = list("transforms[0].value",x),
             label = x)} ) )),
  xaxis = list(range = c(min.x,max.x)),
  yaxis = list(range = c(min.y,max.y)))

p

```

## Error bars

Calculate regression coefficients and standard errors
```{r error_bars_code, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}

m <- lm(data = apra,`Ten-year rate of return` ~ `Operating expense ratio`+ 
          `Investment expenses ratio` + 
          `Proportion of total assets in default or MySuper strategy`)
d <- broom::tidy(m) %>% arrange(desc(estimate))
```

Create a scatter (with the coeffients) and use `error_x` to get the error bars
```{r error_bars_code2, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}
plot_ly(d, x = ~estimate, y = ~term) %>%
  add_markers(error_x = ~list(value = std.error)) %>%
  layout(yaxis = list(title ="", autorange = "reversed")) 

```

## Error bars

```{r error_bars_output, echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE}

m <- lm(data = apra,`Ten-year rate of return` ~ `Operating expense ratio`+ 
          `Investment expenses ratio` + 
          `Proportion of total assets in default or MySuper strategy`)

m2 <- lm(data = apra,`One-year rate of return` ~ `Operating expense ratio`+ 
          `Investment expenses ratio` + 
          `Proportion of total assets in default or MySuper strategy`)

d <- broom::tidy(m) %>% arrange(desc(estimate)) 
d2 <- broom::tidy(m2) %>% arrange(desc(estimate)) %>% mutate(std.error2 = std.error,estimate2=estimate)

d <- left_join(d,d2 %>% select(term,estimate2,std.error2),by = "term")
plot_ly(d, y = ~term) %>%
  add_markers(x = ~estimate, error_x = ~list(value = std.error), name = '10 year rate of return') %>%
  add_markers(x = ~estimate2, error_x = ~list(value = std.error2), name = '1 year rate of return') %>% 
  layout(yaxis = list(title ="", autorange = "reversed"),legend = list(x = 0.7, y = 0.1)) 

```


## Ribbons

```{r ribbons, echo=TRUE, message=FALSE, warning=FALSE, eval=FALSE}

ribs <- lm(`Ten-year rate of return` ~ `Operating expense ratio`, 
           data = apra) %>% augment(.) 

apra$.rownames <- rownames(apra)
apra$.se.fit <- NULL
apra$.fitted <- NULL
apra <- left_join(apra,ribs %>% select(.rownames,.se.fit,.fitted),by=".rownames")

plot_ly(apra, x = ~`Operating expense ratio`, y = ~`Ten-year rate of return`, 
        type = 'scatter', mode = 'markers',text = ~`Fund name`) %>% 
  add_ribbons(
    ymin = ~.fitted - 1.96 * .se.fit,
    ymax = ~.fitted + 1.96 * .se.fit,
    line = list(color = 'rgba(7, 164, 181, 0.05)'),
    fillcolor = 'rgba(7, 164, 181, 0.2)',
    name = "Standard Error") %>% 
  layout(
    xaxis = list(range = c(0, 1)),
    yaxis = list(range = c(-5, 5)),
    showlegend = FALSE)
```

## Ribbons

```{r ribbons_output, echo=FALSE, message=FALSE, warning=FALSE}

ribs <- lm(`Ten-year rate of return` ~ `Operating expense ratio`, 
           data = apra) %>% augment(.) 

apra$.rownames <- rownames(apra)
apra$.se.fit <- NULL
apra$.fitted <- NULL
apra <- left_join(apra,ribs %>% select(.rownames,.se.fit,.fitted),by=".rownames")

plot_ly(apra, x = ~`Operating expense ratio`, y = ~`Ten-year rate of return`, 
        type = 'scatter', mode = 'markers',text = ~`Fund name`) %>% 
  add_ribbons(
    ymin = ~.fitted - 1.96 * .se.fit,
    ymax = ~.fitted + 1.96 * .se.fit,
    line = list(color = 'rgba(7, 164, 181, 0.05)'),
    fillcolor = 'rgba(7, 164, 181, 0.2)',
    name = "Standard Error") %>% 
  layout(
    xaxis = list(range = c(0, 1)),
    yaxis = list(range = c(-5, 5)),
    showlegend = FALSE)
```

## Conclusion

- It is very easy to create impressive and highly informative interative graphics with plotly
- It gets a little more complicated with some features (e.g. subplot,drop down menus,ribbons)
- In some cases, ggploty might be the better way to go (i.e. construct your graph in ggplot then convert to plotly)

**Resources**

- [Carson Sievert's plotly for R](https://plotly-book.cpsievert.me)

- [Plotly R open source graphing library](https://plot.ly/r/)



