---
title: "Access, clean and plot APRA Annual Fund Level Superannuation Statistics - Word version"
author: "Matthew Forbes"
date: "1 November 2018"
output:   
  word_document:
    reference_docx: styles.docx
---

This `R Notebook` contains executable code that will download APRA's Annual Fund Level Superannuation Statistics, and import it into R. The spreadsheet will then be tidied for use in putting together some indicators of superannuation fund outputs.

Note: I have changed all the code chunk settings, to read `echo = FALSE` so as to suppress the inclusion of code in the output.

To produce a Microsoft Word document, you will need to `knit` this document to Word.


## Access data

If the data is not there, the code chunk below will download the data and import Table 2. Table 2 of this dataset provides information about superannuation funds' profile and structure. Table 3 contains information about individual funds' financial performance.

```{r AccessData, echo=FALSE}
# Libraries
library(tidyverse)
library(readxl)
library(pander) 
library(PCcharts)
options(scipen = 999) 

# Download the APRA data if it is not there
if("2018-AFLSS-201706.xlsx" %in% list.files() == FALSE) { 
   link <- "https://www.apra.gov.au/sites/default/files/2018-AFLSS-201706.xlsx"
   download.file(link, "2018-AFLSS-201706.xlsx", mode = "wb")
   }

# Import Table 2
table2 <- read_xlsx("2018-AFLSS-201706.xlsx", sheet = "Table 2", range = "A5:AA206", col_names = TRUE, trim_ws = TRUE) 
table2 <- table2 %>% 
   filter(!is.na(`Fund name`) )
# Remove carriage return and line feed from colnames
colnames(table2) <- gsub("\r\n", " ", colnames(table2) )

# Import Table 3
table3 <- read_xlsx("2018-AFLSS-201706.xlsx", sheet = "Table 3", range = "A5:CE206", col_names = TRUE, trim_ws = TRUE)
table3 <- table3 %>%
   select(`Fund name`, `Total administration and operating expenses`, `Total Investment expenses`, `Cash flow adjusted net assets`) %>% 
   filter(!is.na(`Fund name`) )

# Join tables 2 and 3
apra <- table2 %>% 
   left_join(table3, by = "Fund name")

rm(table2, table3)

```

## Tidy data

There are `r length(apra)` variables for `r nrow(apra)` superannuation funds in our dataset.

To tidy the data for use in this session, we will:

- Select a few variables that are of interest.
- Change the `type` of the variables that we want as numeric from character to numeric. 

```{r TidyData, echo=FALSE, message=FALSE, warning=FALSE}

apra <- apra %>% 
# Select required variables
   select(`Fund name`, `Fund type`, `Total assets`, `Number of MySuper products authorised`, 
          `Total number of member accounts`, `Average member account balance` ,
          `Total administration and operating expenses`, `Total Investment expenses` , `Cash flow adjusted net assets`, 
          `Proportion of total assets in default or MySuper strategy`,
          `One-year rate of return`, `Five-year rate of return`, `Ten-year rate of return` ) %>% 
# Convert vars that we want to be numeric
   mutate_at(.tbl = .,.vars = 3:length(.),.funs = as.numeric ) %>% 
# Turn proportions into percentages
   mutate_at(.tbl = .,.vars = 10:length(.),.funs = funs( .*100) ) %>% 
# Tidy fund type variable
   mutate(`Fund type` = case_when(`Fund type` == "Public Sector" ~ "Public sector" ,
                                `Fund type` == "Retail - ERF" ~ "Retail" , 
                                TRUE ~ `Fund type`) ) %>% 
   mutate(`Fund type` = fct_relevel(as_factor(`Fund type`), c("Industry", "Retail", "Corporate", "Public sector") ) ) %>% 
# Calculate operating and invesetment expenses ratios
   mutate(`Operating expense ratio` = `Total administration and operating expenses`/ `Cash flow adjusted net assets`*100 , 
          `Investment expenses ratio` = `Total Investment expenses`/ `Cash flow adjusted net assets`*100 ) %>% 
# Remove unnecessary variables
   select( -c("Total administration and operating expenses" , "Total Investment expenses" , "Cash flow adjusted net assets") )

# head(apra)
   
```


We can now see that the apra data.frame now has `r length(apra)` variables, and that where data was withheld, it is now `NA`.

## Characteristics of superannuation funds

```{r SummariseData, echo=FALSE, message=FALSE, warning=FALSE}

apra <- apra %>%
   filter(!is.na(`Total assets`) & `Total assets` != 0 ) %>%
   filter(!is.na(`Total number of member accounts`) & `Total number of member accounts` != 0 ) %>% 
# Assets are initially in $'000s
   mutate(`Total assets` = `Total assets`*1000) 

write_csv(apra, "apra.csv")

# Create summary table
summary_table <- apra %>%
   group_by(`Fund type`) %>%
   summarise(`Mean assets ($m)` = sum(`Total assets`, na.rm=TRUE)/n()/1000000 ,
             `Number of funds` = n() ,
             `Operating expense ratio (%)` = weighted.mean(`Operating expense ratio`, `Total assets`, na.rm = TRUE) ,
             `Investment expense ratio (%)` = weighted.mean(`Investment expenses ratio`, `Total assets`, na.rm = TRUE) ,
             `One-year return (%)` = weighted.mean(`One-year rate of return`, `Total assets`, na.rm = TRUE) ,
             `Five-year return (%)` = weighted.mean(`Five-year rate of return`, `Total assets`, na.rm = TRUE) ,
             `Ten-year return (%)` = weighted.mean(`Ten-year rate of return`, `Total assets`, na.rm = TRUE) )

```

The tidied data can then be used to generate an aggregated summary table of the mean characteristics of the different fund types. There are a number of different ways of presenting tables in R markdown:

- When rendering tables into `html`, I find that the `kable` function from the `kableExtra` package provides enough flexibility and produces tables that look quite nice.
- When rendering into word, I prefer `pander::pandoc.table`.

#### Super fund *mean* characteristics
```{r SummaryTable, echo=FALSE, results='asis' }
pandoc.table(summary_table, 
             big.mark = " ", 
             digits = c(0,0,2,2,2,3,3,3),
             justify = c('left', 'right', 'right', 'right','right','right','right','right') ,
             split.tables = Inf)

```

## Let's make some plots

The summary data table above *may* be useful for a report. But it doesn't reallly tell us much about the dataset itself.

It could be useful to look at the returns for individual funds, and then see if funds with higher investment expense ratios (investment expenses as a proportion of total assets) have higher returns.
In order to do this:

1. Insert a new code chunk.
2. Set `echo = FALSE, warning = FALSE` in the header to hide the code chunk. (Note, if you would like to see the code used to generate the plot and any accompanying warnings, set these `=TRUE`.  
3. Write code for a plot. 

### Operating expenses ratio by rate of return 
#### *Circle size indicates funds under mananagment*
```{r OpExpPlot, echo=FALSE, message=FALSE, warning=FALSE}

apra %>%
# Note: Total assets are squared here to emphasise differences in fund size
   mutate(`Total assets` = `Total assets`^2 ) %>% 
   ggplot( aes(x = `Operating expense ratio`, y = `Ten-year rate of return`) ) +
   geom_point( aes(size = `Total assets`, colour = `Fund type`), shape = 1, stroke=1.1) +
   geom_smooth()+
   PC.theme.scat() +
   scale_colour_manual(values = c(PC_green, PC_orange, PC_purple, PC_yellow) ) +
   scale_x_continuous( limits=c(0, 1) ) +
   guides(size=FALSE) 

```

### Investment expenses ratio by rate of return 
#### *Circle size indicates funds under mananagment*
```{r InvExpPlot, echo=FALSE, message=FALSE, warning=FALSE}

apra %>% 
   mutate(`Total assets` = `Total assets`^2 ) %>% 
   ggplot( aes(x = `Investment expenses ratio`, y = `Ten-year rate of return`) ) +
   geom_point( aes(size = `Total assets`, colour = `Fund type`), shape = 1, stroke=1.1) +
   geom_smooth()+
   PC.theme.scat() +
   scale_colour_manual(values = c(PC_green, PC_orange, PC_purple, PC_yellow) ) +
   scale_x_continuous( limits=c(0, 1) ) +
   guides(size=FALSE) 

```

### Investment and operating expenses ratios
#### *Circle size indicates funds under mananagment*
```{r InvestOpExpPlot, echo=FALSE, message=FALSE, warning=FALSE}

apra %>% 
   mutate(`Total assets` = `Total assets`^2 ) %>% 
   ggplot( aes(y = `Investment expenses ratio`, x = `Operating expense ratio`, colour = `Fund type`) ) +
   geom_point( aes(size = `Total assets` ), shape = 1, stroke=1.5 ) +
   scale_colour_manual(values = c(PC_green, PC_orange, PC_purple, PC_yellow) ) +
   scale_y_continuous( limits=c(0,1.5) ) +
   scale_x_continuous( limits=c(0,5) ) +
   PC.theme.scat()+
   guides(size=FALSE) 

```

### Returns by proportion of default/MySuper assets
#### *Circle size indicates funds under mananagment*
```{r ReturnsMySuperPlot, echo=FALSE, message=FALSE, warning=FALSE}

apra %>% 
   mutate(`Total assets` = `Total assets`^2 ) %>% 
   ggplot( aes(y = `Ten-year rate of return`, x = `Proportion of total assets in default or MySuper strategy`) ) +
   geom_point( aes(size = `Total assets` , colour = `Fund type` ), shape = 1, stroke = 1.05) +
   scale_colour_manual(values = c(PC_green, PC_orange, PC_purple, PC_yellow) ) +
   scale_y_continuous( limits=c(-2,7) ) +
   geom_smooth()+
   PC.theme.scat()+
   guides(size=FALSE) 

```
